Refinement for open automata

### 2021-06-08 2021-Refinement/report.tex
Sync

For three open automata \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3 \defobject \OAg[3]\), such that \(R_{12}\) is a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) and \(R_{23}\) of \(A_2\) by \(A_3\), we want to define \(R_{13}\) by saying that there exists a state \(s_2 \in S_2\) and values for \(V_2\) such that \(R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}\).
However we cannot talk about states in formulas.
The solution is to expand the existential quantifier as we did in deadlock prevention, which is possible because states are finite.
\nmm{R_{13} \defobject \mpar{s_1, s_3} \mapsto \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}} is a witness of the transitivity, proof is given in appendix \ref{apx:trans}.
\begin{thm}[Context refinement] Hole-equal simulation is context refining \end{thm}
\begin{proof}
% TODO: assume a <= b & a[c] -> b[c] + proof
\end{proof}
% Rédiger un paragraphe sur le fait que composer dans l'implem sans deadlock implique qu'il n'y a pas de deadlock quand on compose avec la spec, on voudrais avoirs l'autre sens mais ce n'est pas possible
\begin{prop}[Composition correctness'] Hole-subset simulation is correct w.r.t.\@ composition. \end{prop}
\begin{proof}
% TODO
\end{proof}
\begin{prop}[Composition correctness"]
Let \(A_1 \defobject \OAg[1]\) be hole-superset simulated by \(A_2 \defobject \OAg[2]\).
Let \(j \in J_1 \setminus J_2\) and \(A_3\) an open automaton.
There is a hole-superset simulation of \(A_1\subst{A_3}{j}\) by \(A_2\).
\end{prop}
\begin{proof}
% TODO
\end{proof}
% J'2 \subseteq J'1 in hole-subset, J'1 \subseteq J'2 in hole-superset (decide of transition variable with holes); composition with new holes is hole-subset where hole variables are free then hole-superset where hole names have no conflicts with removed holes. Transitivity must ensure holes name have no conflict for that reason!

@@

\section{Flat model}
% Goal: some interfaces refine in one way while other refine in the other way and others are equivalent and some are untracked.

@@

	\(R\) satisfies \eqref{eq:drWD} with the current value of \(\mpar{s_1, s_2}\) and \(\sigma \uplus \nu\) completed with dummy values for the other variables of \(\biguplus_{t_2 \in \fOT{s_2}} \fvars{t_2}\) so we have:

@@

	In order to get a \(t_2\) matching \(t_1\) we will use the fact that \(R\) is a simulation.

@@

	\sigma_{01} \uplus \sigma_{03} \vdash R_{13}\mpar{s_{01}, s_{03}} \\
	\equiv \\
	\sigma_{01} \uplus \sigma_{03} \vdash \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_{01}, s_2} \wedge R_{23}\mpar{s_2, s_{03}} \\
	\Uparrow \\
	\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03} \vdash R_{12}\mpar{s_{01}, s_{02}} \wedge R_{23}\mpar{s_{02}, s_{03}} \\
	\Uparrow \\
	\mpar{\sigma_{01} \uplus \sigma_{02} \vdash R_{12}\mpar{s_{01}, s_{02}}} \wedge \mpar{\sigma_{02} \uplus \sigma_{03} \vdash R_{23}\mpar{s_{02}, s_{03}}}

@@

		\bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{t_{3x} \defobject \OTx{3}{x}{3x}{3x} \in \fOT{s_3}}^{x \in X}, \\
		\mpar{\forall x \in X, J'_{3x} \cap J_1 = J'_1 \cap J_3} \\
		\wedge V_1 \uplus V_3 \uplus \fvars{t_1} \vdash R_{13}\mpar{s_1, s_3} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}\begin{array}{l}

@@

	\begin{multline*}
		\mpar{R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}}\psubst{\psi_1 \uplus \psi_{3z}}\psubst{\psi_{s2x}\psubst{\sigma_2 \uplus \nu_2}}\psubst{\sigma_1 \uplus \sigma_3 \uplus \nu_3} \\
		= R_{12}\mpar{s_1, s_2}\psubst{\psi_1}\psubst{\psi_{s2x}\psubst{\sigma_2 \uplus \nu_2}}\psubst{\sigma_1} \wedge R_{23}\mpar{s_2, s_3}\psubst{\psi_{3z}}\psubst{\psi_{s2x}\psubst{\sigma_2 \uplus \nu_2}}\psubst{\sigma_3 \uplus \nu_3} \\
		= R_{12}\mpar{s_1, s_2}\psubst{\psi_1 \uplus \psi_{s2x}}\psubst{\sigma_2 \uplus \nu_2 \uplus \sigma_1} \wedge R_{23}\mpar{s_2, s_3}\psubst{\psi_{3z} \uplus \psi_{s2x}}\psubst{\sigma_2 \uplus \nu_2 \uplus \sigma_3 \uplus \nu_3}
	\end{multline*}

>>>>

For three open automata \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3 \defobject \OAg[3]\), such that \(R_{12}\) is a hole-\(\triangle\) simulation of \(A_1\) by \(A_2\) and \(R_{23}\) of \(A_2\) by \(A_3\), we want to define \(R_{13}\) by saying that there exists a state \(s_2 \in S_2\) and values for \(V_2\) such that \(R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}\).
However we cannot talk about states in formulas.
The solution is to expand the existential quantifier on states as we did in deadlock prevention, which is possible because states are finite.
\nmm{R_{13} \defobject \mpar{s_1, s_3} \mapsto \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}} is a witness of the transitivity, proof is given in appendix \ref{apx:trans}.
\begin{thm}[Context refinement] Hole-equal simulation is context refining \end{thm}
	For three open automata \(A_1 \defobject \OAg[1]\), \(A_2 \defobject \OAg[2]\) and \(A_3 \defobject \OAg[3]\), let \(k \in J_1\), \(A_{13} \defobject A_1\subst{A_3}{k}\) and \(A_{23} \defobject A_2\subst{A_3}{k}\) with \(A_{13} = \OAg[13]\) and \(A_{23} = \OAg[23]\).
	Let \(R_{12}\) be a hole-equal simulation of \(A_1\) by \(A_2\) and \(\reach{A_{13}}\) be a witness that \(A_{13}\) is non-locking.
	The idea for the relation on configurations is that \(A_2\) simulates \(A_1\) while \(A_3\) does the same things in both compositions.
	We pose
	\[ R \defobject \mpar{\mpar{s_1, s_{31}}, \mpar{s_2, s_{32}}} \mapsto \choice{R_{12}\mpar{s_1, s_2} \wedge \bigwedge_{v_3 \in V_3} v_{31} = v_{32} & \text{if } s_{31} = s_{32} \\ \bot & \text{otherwise}} \]
	Where \(y_{3x}\) is the renaming of \(y_3\) from \(A_3\) in \(A_{x3}\).
	We want to prove that \(R\) is a hole-equal simulation of \(A_{12}\) by \(A_{23}\):
		J_1 \cup J_3 \setminus \mbrc{k} = J_2 \cup J_3 \setminus \mbrc{k} \goal{1a} \\
		\sigma_{013} \uplus \sigma_{023} \vdash R\mpar{s_{013}, s_{023}} \goal{1b}
	\end{gather}\vspace{-1cm}
		\forall \mpar{s_{13}, s_{23}} \in S_{13} \times S_{23}, \\
			\bigsymb{\forall} t_{13} \defobject \OTx{13}{}{13}{13} \in \fOT{s_{13}}, \bigsymb{\exists} \mpar{t_{23x} \defobject \OTx{23}{x}{23x}{23x} \in \fOT{s_{23}}}^{x \in X}, \\
			\quad \mpar{\forall x \in X, J'_{23x} \cap J_{13} = J'_{13} \cap J_{23}} \\[-10pt]
			\nwedge V_{13} \uplus V_{23} \uplus \fvars{t_{13}} \vdash R\mpar{s_{13}, s_{232}} \wedge g_{13} \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\begin{array}{l}
				\alpha_{13} = \alpha_{23x} \wedge \bigwedge_\subbox{j \in J'_{23x} \cap J_{13}} \beta_{13j} = \beta_{23xj} \\[12pt]
				\nwedge g_{23x} \wedge R\mpar{s'_{13}, s'_{23x}}\psubst{\psi_{13} \uplus \psi_{23x}}
		\end{array}} \\
		\wedge V_{13} \uplus V_{23} \uplus \biguplus_\subbox{t_{23} \in \fOT{s_{23}}} \fvars{t_{23}} \vdash R\mpar{s_{13}, s_{23}} \wedge \bigvee_\subbox{t_{23} \in \fOT{s_{23}}} \fguard{t_{23}} \implies \bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}} \goal{1c}
	\goal{1a} is proved by rewriting using the hypothesis \(J_1 = J_2\) from the simulation.
	For \goal{1b} we use the fact that \(R_{12}\) relates inital configurations:
		\sigma_{013} \uplus \sigma_{023} \vdash R\mpar{s_{013}, s_{023}} \iff & R\mpar{\mpar{s_{01}, s_{031}}, \mpar{s_{02}, s_{031}}}\psubst{\sigma_{01} \uplus \sigma_{02} \uplus \sigma_{031} \uplus \sigma_{032}} \\
		\iff & R_{12}\mpar{s_{01}, s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02}} \wedge \mpar{\bigwedge_{v_3 \in V_3} v_{31} = v_{32}}\psubst{\sigma_{031} \uplus \sigma_{032}} \\
		\iff & R_{12}\mpar{s_{01}, s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02}} \wedge \bigwedge_{v_3 \in V_3} \sigma_{031}\mpar{v_{31}} = \sigma_{032}\mpar{v_{32}} \\
		\iff & R_{12}\mpar{s_{01}, s_{02}}\psubst{\sigma_{01} \uplus \sigma_{02}} \wedge \top \\
		\iff & \sigma_{01} \uplus \sigma_{02} \vdash R_{12}\mpar{s_{01}, s_{02}}
	Let \(s_{13} \defobject \mpar{s_1, s_{31}} \in S_{13}\) and \(s_{23} \defobject \mpar{s_2, s_{32}} \in S_{23}\).
	The goals left to prove are
		& \bigsymb{\forall} t_{13} \defobject \OTx{13}{}{13}{13} \in \fOT{s_{13}}, \bigsymb{\exists} \mpar{t_{23x} \defobject \OTx{23}{x}{23x}{23x} \in \fOT{s_{23}}}^{x \in X}, \\
		& \quad \mpar{\forall x \in X, J'_{23x} \cap J_{13} = J'_{13} \cap J_{23}} \\[-10pt]
		& \nwedge V_{13} \uplus V_{23} \uplus \fvars{t_{13}} \vdash R\mpar{s_{13}, s_{232}} \wedge g_{13} \Rightarrow \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\!\!\everymath{\displaystyle}\begin{array}{l}
			\alpha_{13} = \alpha_{23x} \wedge \bigwedge_\subbox{j \in J'_{23x} \cap J_{13}} \beta_{13j} = \beta_{23xj} \\[12pt]
			\nwedge g_{23x} \wedge R\mpar{s'_{13}, s'_{23x}}\psubst{\psi_{13} \uplus \psi_{23x}}
		\end{array}\!\!} \goal{2}
	\[ V_{13} \uplus V_{23} \uplus \biguplus_\subbox{t_{23} \in \fOT{s_{23}}} \fvars{t_{23}} \vdash R\mpar{s_{13}, s_{23}} \wedge \bigvee_\subbox{t_{23} \in \fOT{s_{23}}} \fguard{t_{23}} \implies \bigvee_\subbox{t_{13} \in \fOT{s_{13}}} \fguard{t_{13}} \goal{3} \]
\item[\goal{2}:] % TODO: introduce t_... then hyps
\item[\goal{3}:] % TODO: introduce hyps
% Rédiger un paragraphe sur le fait que composer dans l'implem sans deadlock implique qu'il n'y a pas de deadlock quand on compose avec la spec, on voudrais avoirs l'autre sens mais ce n'est pas possible
\begin{prop}[Composition correctness'] Hole-subset simulation is correct w.r.t.\@ composition. \end{prop}
	Let \(A_1 \defobject \OAg[1]\) and \(A_2 \defobject \OAg[2]\) be open automata.
	Let \(k \in J_1\) and \(A \defobject A_1\subst{A_2}{k}\).
	Let \(\reach{A}\) be a witness that \(A\) is non-locking.
	The idea for the relation on configurations is that the copy of \(A_1\) in \(A\) do the same things as if we were to try proving reflexivity.
	We pose
	\[ R \defobject \mpar{\mpar{s_1, s_2}, s'_1} \mapsto \choice{\bigwedge_{v \in V_1} v = v' & \text{if } s_1 = s'_1 \\ \bot & \text{otherwise}} \]
	% TODO: proove A \leq A_1
\begin{prop}[Composition correctness"]
Let \(A_1 \defobject \OAg[1]\) be hole-superset simulated by \(A_2 \defobject \OAg[2]\).
Let \(j \in J_1 \setminus J_2\) and \(A_3\) an open automaton.
There is a hole-superset simulation of \(A_1\subst{A_3}{j}\) by \(A_2\).
\end{prop}
% TODO
% J'2 \subseteq J'1 in hole-subset, J'1 \subseteq J'2 in hole-superset (decide of transition variable with holes); composition with new holes is hole-subset where hole variables are free then hole-superset where hole names have no conflicts with removed holes. Transitivity must ensure holes name have no conflict for that reason!

@@

@@

	\(R\) satisfies \eqref{eq:drWD} with the current value of \(\mpar{s_1, s_2}\) and \(\sigma \uplus \nu\) completed with dummy values for the other variables of \nmm{\biguplus_{t_2 \in \fOT{s_2}} \fvars{t_2}} so we have:

@@

	In order to get a \(t_2\) matching \(t_1\) we will use the fact that \(R\) is a pre-simulation.

@@

	\sigma_{01} \uplus \sigma_{03} \vdash R_{13}\mpar{s_{01}, s_{03}} \iff & \sigma_{01} \uplus \sigma_{03} \vdash \exists V_2, \bigvee_\subbox{s_2 \in S_2} R_{12}\mpar{s_{01}, s_2} \wedge R_{23}\mpar{s_2, s_{03}} \\
	\impliedby & \sigma_{01} \uplus \sigma_{02} \uplus \sigma_{03} \vdash R_{12}\mpar{s_{01}, s_{02}} \wedge R_{23}\mpar{s_{02}, s_{03}} \\
	\impliedby & \mpar{\sigma_{01} \uplus \sigma_{02} \vdash R_{12}\mpar{s_{01}, s_{02}}} \wedge \mpar{\sigma_{02} \uplus \sigma_{03} \vdash R_{23}\mpar{s_{02}, s_{03}}}

@@

		& \bigsymb{\forall} t_1 \defobject \OTx{1}{}{1}{1} \in \fOT{s_1}, \bigsymb{\exists} \mpar{t_{3x} \defobject \OTx{3}{x}{3x}{3x} \in \fOT{s_3}}^{x \in X}, \\
		& \mpar{\forall x \in X, J'_{3x} \cap J_1 = J'_1 \cap J_3} \\[-10pt]
		& \wedge V_1 \uplus V_3 \uplus \fvars{t_1} \vdash R_{13}\mpar{s_1, s_3} \wedge g_1 \implies \operatorname*{\bigsymb{\bigvee}}_{x \in X} \mpar{\everymath{\displaystyle}\begin{array}{l}

@@

		& \sigma_1 \uplus \sigma_3 \uplus \nu_3 \vdash R_{13}\mpar{s'_1, s'_{3z}}\psubst{\psi_1 \uplus \psi_{3z}} \\
		\iff & \mpar{R_{12}\mpar{s_1, s_2} \wedge R_{23}\mpar{s_2, s_3}}\psubst{\psi_1 \uplus \psi_{3z}}\psubst{\psi_{s2x}\psubst{\sigma_2 \uplus \nu_2}}\psubst{\sigma_1 \uplus \sigma_3 \uplus \nu_3} \\
		\iff & R_{12}\mpar{s_1, s_2}\psubst{\psi_1}\psubst{\psi_{s2x}\psubst{\sigma_2 \uplus \nu_2}}\psubst{\sigma_1} \wedge R_{23}\mpar{s_2, s_3}\psubst{\psi_{3z}}\psubst{\psi_{s2x}\psubst{\sigma_2 \uplus \nu_2}}\psubst{\sigma_3 \uplus \nu_3} \\
		\iff & R_{12}\mpar{s_1, s_2}\psubst{\psi_1 \uplus \psi_{s2x}}\psubst{\sigma_2 \uplus \nu_2 \uplus \sigma_1} \wedge R_{23}\mpar{s_2, s_3}\psubst{\psi_{3z} \uplus \psi_{s2x}}\psubst{\sigma_2 \uplus \nu_2 \uplus \sigma_3 \uplus \nu_3} \\
		\iff & \mpar{\sigma_2 \uplus \nu_2 \uplus \sigma_1 \vdash R_{12}\mpar{s_1, s_2}\psubst{\psi_1 \uplus \psi_{s2x}}} \wedge \mpar{\sigma_2 \uplus \nu_2 \uplus \sigma_3 \uplus \nu_3 \vdash R_{23}\mpar{s_2, s_3}\psubst{\psi_{3z} \uplus \psi_{s2x}}}
